### 2013-04-01 writeup/motivation.tex
cleaning

\section{Why Availability?}
\label{sec:motivation}

Why does availability matter? Peter Deutsch starts his classic list of
``Fallacies of Distributed Computing'' with two concerns fundamental
to distributed database systems: ``\textit{1.)}  The network is
reliable. \textit{2.)} Latency is zero''~\cite{fallacies-deutsch}; in
a distributed setting, network failures may prevent database servers
from communicating, and, in the absence of failures, communication is
slowed by factors like physical distance and, more often, congestion
and routing. As we will see (Section~\ref{sec:availability}), highly
available systems minimize the effect of these network
\textit{partitions} and \textit{latency}. In this Section, we draw on
a range of evidence that indicates that partitions occur relatively
frequently in real-world deployments and best-case latencies between
datacenters are significant.

\subsection{Network Partitions at Scale}

According to James Hamilton, Vice President and Distinguished Engineer
on the Amazon Web Services team, ``network partitions should be rare
but net gear continues to cause more issues than it
should''~\cite{hamilton-partitions}. Anecdotal evidence confirms
Hamilton's assertion. In April 2011, a network misconfiguration led to
a twelve-hour long series of outages across the Amazon EC2 and RDS
services~\cite{amazon-netpartition}. Subsequent misconfigurations and
partial failures such as another EC2 outage in October 2012 have led
to full site disruptions for popular web services like Reddit,
Foursquare, and Heroku~\cite{ec2-downsites}. At a global scale,
hardware failures---like the 2011 outages in Internet backbones in
North America and Europe due a router
bug~\cite{juniper-partition}---and misconfigurations such as BGP
faults in 2008~\cite{pakistan-youtube} and
2010~\cite{research-experiment-partition} can cause widespread
partitions.

Many of our discussions with practitioners---especially those
operating on public cloud infrastructure---as well as reports from
large-scale operators like Google~\cite{dean-keynote} confirm that
partition management is an important consideration for service operators
today. As an example of an actual system rearchitecture resulting from
partitions, less than one year after its announcement, Yahoo!'s PNUTS
developers explicitly added support for weaker, highly available
operation. The engineers explained that ``strict adherence [to strong
  consistency] leads to difficult situations under network
partitioning or server failures...in many circumstances, applications
need a relaxed approach''~\cite{pnuts-update}.

Several recent studies rigorously quantify partition behavior. A 2011
study of several Microsoft datacenters found a mean of 40.8 network
link failures per day (95th percentile: 136), with a median time to
repair of around five minutes (and up to one week). Perhaps
surprisingly, provisioning redundant networks only reduces impact of
failures by up to 40\%, meaning network providers cannot easily
curtail partition behavior~\cite{sigcomm-dc}. A 2010 study of over 200
wide-area routers found an average of 16.2--302.0 failures per link
per year with an average annual downtime of 24--497 minutes per link
per year (95th percentile at least 34 hours)~\cite{sigcomm-wan}. In
HP's managed enterprise networks, WAN, LAN, and connectivity problems
account for 28.1\% of all customer support tickets while 39\% of
tickets relate to network hardware.  The median incident duration for
highest priority tickets ranges from 114--188 minutes and up to a full
day for all tickets~\cite{turner2012failure}. Other studies confirm
these results, showing median time between failures over a WAN network
of approximately 3000 seconds with a median time to repair between 2
and 1000 seconds~\cite{ip-backbone-failures} as well as frequent path
routing failures on the Internet~\cite{labovitz-failures}. Isolating,
quantifying, and designing for these network failures is an area of
active research in networking
community~\cite{uw-failure-networks}.

These results---which do not even consider server-level
failures---indicate that network partitions \textit{do} occur within
and across modern datacenters. We do not intend to fixate on any
particular statistical behavior but instead observe that, as a general
trend, partitions exist and must accordingly be met with either
unavailability at some servers or, as we will discuss, relaxed
semantic guarantees.

\subsection{Latency and Planet Earth}
\label{sec:latency}

Even with fault-free networks, distributed systems face the challenge
of communication latency, Deutsch's second ``Fallacy.'' In this
section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds---in a geo-replicated context.
Fundamentally, the speed at which two servers can communicate is
(according to modern physics) bounded by the speed of light. In the
best case, two servers on opposite sides of the Earth's
surface---communicating via a hypothetical link traversing the
planet's diamater---require a minimum 85.1 ms RTT (133.7 ms if sent at
ground level). As services are replicated to multiple, geographically
distinct sites, the cost of communication increases to this lower
bound.

\definecolor{min-lat-color}{HTML}{B2FF99}
\definecolor{max-lat-color}{HTML}{FF7F7F}

\begin{table}
\vspace{.5em}

\subfloat[Across \texttt{us-east} AZs]{
  \makebox[.2\textwidth]{
    \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{C} & \multicolumn{1}{c|}{D}\\\hline
B & \colorbox{min-lat-color}{1.08} & 3.12 \\
C & & \colorbox{max-lat-color}{3.57}  \\
\hline
  \end{tabular}}
 }
\subfloat[Within \texttt{us-east-a} AZ] {
  \makebox[.25\textwidth]{
  \begin{tabular}{|c|c|c|}\hline
 & \multicolumn{1}{c}{H2} & \multicolumn{1}{c|}{H3}\\\hline
H1  & 0.55   & \colorbox{max-lat-color}{0.56} \\
H2 &  & \colorbox{min-lat-color}{0.50}  \\
\hline
  \end{tabular}
 }}

\subfloat[Cross-region (CA:~California, OR:~Oregon, VA:~Virginia, TO:~Tokyo, IR:~Ireland, SY:~Sydney, SP:~S\~{a}o Paulo, SI:~Singapore)] {
  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
& \multicolumn{1}{c}{OR} & \multicolumn{1}{c}{VA} & \multicolumn{1}{c}{TO} & \multicolumn{1}{c}{IR} & \multicolumn{1}{c}{SY} & \multicolumn{1}{c}{SP} & \multicolumn{1}{c|}{SI} \\\hline
CA & \colorbox{min-lat-color}{22.5}   & 84.5   & 143.7   & 169.8   & 179.1   & 185.9   & 186.9  \\
OR &  & 82.9   & 135.1   & 170.6   & 200.6   & 207.8   & 234.4  \\
VA & &  & 202.4   & 107.9   & 265.6   & 163.4   & 253.5  \\
TO & & &  & 278.3   & 144.2   & 301.4   & 90.6  \\
IR & & & &  & 346.2   & 239.8   & 234.1  \\
SY & & & & &  & 333.6   & 243.1  \\
SP & & & & & &  & \colorbox{max-lat-color}{362.8}  \\
\hline
  \end{tabular}
}

\caption{Mean RTT times on EC2 (min and max highlighted)}
\label{table:rtt}
\end{table}

\begin{figure}
\includegraphics[width=\columnwidth]{graphs/ping-plot.pdf}
\caption{CDF of round-trip times for slowest inter- and intra-
  availability zone links compared to cross-region links.}
\label{fig:rtt}
\end{figure}

In real deployments, transit times are actually slower than the speed
of light due to routing, congestion, and processing times. To
illustrate the difference between intra-datacenter, inter-datacenter,
and inter-planetary networks, we performed a measurement study of
network behavior on Amazon's EC2, a widely-used public compute
cloud. We measured one week of ping times between all seven EC2
geographic ``regions'', across three ``availability zones'' (closely
co-located datacenters), and within a single ``availabilty zone''
(datacenter), at a granularity of 1s (dataset to be released). We
summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, intra-datacenter communication (1a)
is between 1.82 and 6.38 times faster than across geographically
co-located datacenters (1b) and between 40 and 647 times faster than
across geographically distributed datacenters (1c). The cost of
wide-area communication exceeds the speed of light: for example,
communicating from S\~{a}o Paulo to Singapore is lower-bounded by
speed-of-light transit RTT of 106.7ms but, on average, incurs a
362.8ms RTT (95th percentile: 649ms). As shown in
Figure~\ref{fig:rtt}, the distribution of latencies varies between
links, but the trend is clear: remote communication has a substantial
cost.

\section{When is ``ACID'' ACID?}
\label{sec:modernacid}

The previous section demonstrated that distributed systems must
address partitions and latency: what does this mean for distributed
databases? Database researchers and designers have long realized that
serializability is not achievable in a highly available
system~\cite{davidson-survey}, meaning it will be expensive in the
environments from Section~\ref{sec:motivation}. However, even in a
single-node database, the coordination penalties associated with
serializability can be severe, manifesting themselves in the form of
decreased concurrency (and, subsequently, performance degradation,
scalability limitations, and, often, external aborts like
deadlocks)~\cite{gray-isolation}. Accordingly, to increase concurrency,
database systems offer a range of ACID properties weaker than
serializability.  Instead, databases offer a host of so-called
\textit{weak isolation} models that allow varying restrictions on the
space of schedules that are allowable by the system~\cite{adya,
  ansicritique, ansi-sql}. None of these weak isolation models
guarantees serializability, but, as we see below, the benefits of
these models are often considered to outweigh costs of possible
consistency anomalies that might arise from their use.

To understand the prevalence of these weak isolation models, we
recently surveyed the default and maximum isolation models provided by
18 databases, often claiming to provide ``ACID'' or ``NewSQL''
functionality~\cite{hat-hotos}. As shown in
Table~\ref{table:existing}, only three out of 18 databases provided
serializability by default, and eight did not provide serializability
as an option at all. This is particularly surprising when we consider
the widespread deployment of many of these non-serializable databases,
like Oracle 11g, which are known to power major businesses and product
functionality. Given that these transactional models are frequently
used, our inability to provide serializability in arbitrary HATs
appears non-fatal for practical applications. If application writers
and database vendors have already decided that the benefits of weak
isolation outweigh potential application inconsistencies, then, in a
highly available environment that prohibits serializability, similar
decisions may be tenable.

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{|l|c|c|}
\hline
Database & Default & Maximum\\\hline
Actian Ingres 10.0/10S & S & S\\
Aerospike & RC & RC\\
Akiban Persistit & SI & SI\\
Clustrix CLX 4100 & RR & RR\\
Greenplum 4.1 & RC & S \\
IBM DB2 10 for z/OS & CS & S\\
IBM Informix 11.50 & Depends & S\\
MySQL 5.6 & RR & S \\
MemSQL 1b & RC & RC\\
MS SQL Server 2012 & RC & S \\
NuoDB & CR & CR\\
Oracle 11g & RC & SI\\
Oracle Berkeley DB & S & S\\
Oracle Berkeley DB JE & RR & S\\
Postgres 9.2.2 & RC & S\\
SAP HANA & RC & SI\\
ScaleDB 1.02 & RC & RC\\
VoltDB & S & S\\
\hline
\multicolumn{3}{|p{7cm}|}{{\footnotesize RC: read committed, RR: repeatable read, SI: snapshot isolation, S: serializability, CS: cursor stability, CR: consistent read}}\\\hline

\end{tabular}
\caption{Default and maximum isolation levels for ACID and NewSQL
  databases as of January 2013 (from
  \protect\cite{hat-hotos}).}\vspace{-1.5em}
\label{table:existing}
\end{small}
\end{center}
\end{table}

It is unknown \textit{which} of these ACID (particularly isolation)
guarantees can be provided with high availability. The existing
algorithms for providing weak isolation are typically designed for a
single-node context and are, to the best of our knowledge, unavailable
due to a reliance on concurrency control mechanisms like locking that
are not resilient to partial failure
(Section~\ref{sec:evaluation}). Moreover, we are not aware of any prior
literature that provides guidance as to the relationship between weak
isolation and high availability: prior work has examined the
relationship between serializability and high
availability~\cite{davidson-survey} and weak isolation on a single
server~\cite{adya, ansicritique, gray-isolation} but never weak
isolation and high availability taken together.  A primary goal in the
remainder of this paper is to understand which models are
HAT-compliant.