### 2013-04-25 writeup/futurework.tex

\section{Future Work}
\label{sec:futurework}

There are several promising avenues for future work in HATs:

\textbf{What are the limits of HAT semantics?}  In this work, we have
taxonomized several well-defined and well-documented semantic
properties. This is useful as it contextualizes existing database and
distributed systems semantics but does not provide guidance as to what
consistency models are the ``strongest''---or admit the smallest set
of executions---within each availability class. A recent report from
UT Austin shows that no model stronger than causal consistency can be
achieved in a sticky highly available, one-way convergent
system~\cite{cac}, but there are infinite highly available and sticky
highly available models to consider. For example, always returning the
value $1$ in response to reads is incomparable with (i.e., neither
stronger nor weaker than) causal consistency: returning $1$ does not
respect causality, and causal consistency returns values other than
$1$. There are actually infintely many incomparable models (vacuously:
return $1$, return $2$, and so on). Are there \textit{meaningful}
properties that delineate the boundaries between availability levels?
Is there a ``meta-property'' that exactly characterizes what
properties are achievable with high availability or sticky high
availability? This is an area of ongoing research within the Berkeley
HAT project. \pbnote{SHORTEN}

\textbf{What are the costs and benefits within HAT designs?} There are
many possible combinations of HAT guarantees, and the performance and
ease of implementation of a HAT system will vary depending on which
guarantees it provides. For example, as we have discussed, the cost in
moving from a sticky high available implementation of highly available
guarantees to a truly highly available system seems expensive for some
models (e.g., metadata requirements or visibility penalties may
become problematic). Further experimentation regarding the set of
useful and achievable HAT guarantees will shed further light on this
question.

\textbf{Where do semantics-based concurrency control and weakened
  failure models lie in the HAT landscape?} The vast literature on
semantics-based concurrency control indicates when coordination can be
avoided while providing ``consistent'' operation for limited failure
scenarios. We have performed a narrow case study, but a general
taxonomization of techniques such as escrow
transactions~\cite{escrow}, Sagas~\cite{sagas}, and other long-running
transaction support could help address the challenge of programming in
a weakly consistent environment. HAT guarantees can also be
strengthened by considering weaker failure models such as networks
with bounded asynchrony. Additionally, while we have focused on
systems that provide a single availability mode, there are several
interesting hybrids to consider. For example, a system may provide
strong semantics in the absence of partitions, but, in the presence of
partitions, ``lose'' guarantees by having clients reconnect to an
available replica. This provides ``best effort'' guarantees but, to
provide availability, sacrifices them in the event of partitions.