### 2013-05-01 writeup/hat-vldb.tex

% THIS IS AN EXAMPLE DOCUMENT FOR VLDB 2012
% based on ACM SIGPROC-SP.TEX VERSION 2.7
% Modified by  Gerald Weber <gerald@cs.auckland.ac.nz>
% Removed the requirement to include *bbl file in here. (AhmetSacan, Sep2012)
% Fixed the equation on page 3 to prevent line overflow. (AhmetSacan, Sep2012)

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}}

\interfootnotelinepenalty=10000

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, amsthm, float}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\newcommand{\subparagraph}{}

\usepackage{xspace, titlesec}

\newcommand{\etal}{{\em et al.}\xspace}
\newcommand{\etc}{{\em etc}\xspace}
\newcommand{\cf}{{\em c.f.,}\xspace}
\newcommand{\versus}{{\em vs.}\xspace}
\newcommand{\ie}{{\em i.e.,}\xspace}
\newcommand{\eg}{{\em e.g.,}\xspace}

\newcommand{\figfactor}{1.0}

\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx
\newcommand{\commentt}[1]{{\small\texttt{#1}}}

\theoremstyle{definition}
\newtheorem{observation}{Observation}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

\renewcommand{\ttdefault}{fi4}

\usetikzlibrary{positioning, snakes, shapes}

\tikzset{>=latex}

\titleformat*{\section}{\fontsize{12pt}{12}\selectfont\bfseries}
\titleformat*{\subsection}{\fontsize{10pt}{11}\selectfont\bfseries}
\titleformat*{\subsubsection}{\fontsize{10pt}{11}\itshape}

\titlespacing{\section}{1ex}{2ex}{1ex}

\titlespacing{\subsection}{1ex}{1ex}{0.25ex}
\titlespacing{\subsubsection}{1ex}{1ex}{0.25ex}

\let\footnotesize\small

\newcommand*{\DashedArrow}[1][]{\mathbin{\tikz [baseline=-.25ex,-latex, dashed,#1] \draw [#1] (0pt,0.5ex) -- (2em,0.5ex);}}%

\newenvironment{myitemize}
{

   \vspace{0mm}
    \begin{list}{$\bullet$}{\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

\newenvironment{myenumerate}
{

   \vspace{0mm}
   \newcounter{qcounter}
    \begin{list}{\arabic{qcounter}.~}{\usecounter{qcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

\begin{document}

% ****************** TITLE ****************************************

\title{Highly Available Transactions: Virtues and Limitations}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

\begin{abstract}
\noindent To minimize network latency and remain online during server
failures and network partitions, many modern distributed data storage
systems eschew transactional functionality, which provides strong
semantic guarantees for groups of multiple operations over multiple
data items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak isolation, replica consistency, and highly
available systems. We analytically and experimentally quantify the
availability and performance benefits of HATs---often two to three
orders of magnitude over wide-area networks---as well as their
necessary semantic compromises.
\end{abstract}\vspace{-.5em}

\input{intro.tex}

\input{motivation.tex}

\input{model.tex}

\input{hat-analysis.tex}

\input{evaluation.tex}

\input{relatedwork.tex}

%\input{futurework.tex}

\input{conclusion.tex}

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
%\footnotesize
\scriptsize

\bibliography{hat-vldb}

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
%\pagebreak
%
\pagebreak
%\begin{appendix}
%\input{availability-appendix.tex}

\input{hat-appendix.tex}
%\end{appendix}
\end{document}

% megastore
% redblue
% gemini
% redblue
% granola
% bernstein-concurrency
% surviving-failures-bodik
% blooml
% google-availability
% cops