### 2013-05-01 writeup/intro.tex

\section{Introduction}

The last decade has seen a significant shift in the design of
large-scale database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. The 2000 introduction of
Brewer's CAP Theorem, which stated that a highly available system
cannot provide strong consistency guarantees in the presence of
network partitions, helped spur this shift~\cite{brewer-slides}. As
formally proven~\cite{gilbert-cap}, the CAP Theorem concerns
linearizability: the ability to read the most recent write to a data
item that is replicated across servers~\cite{herlihy-art}. Yet the
implications of CAP---which concern a fairly narrow form of
distributed consistency---have often been conflated with the ability
to provide ACID database properties~\cite{hn,brewer-slides,
  foundation-article}; this misunderstanding has led to substantial
confusion regarding replica consistency, transactional isolation, and
high availability. The recent resurgence of transactional systems like
Spanner~\cite{spanner} suggests that programmers value these
semantics, but, unfortunately, existing transactional data stores do
not provide availability in the presence of
partitions~\cite{middleware-db, foundation-article,
  hstore,spanner,generalizedsnapshot, mdcc, krikellas-bargain, eiger,
  walter, calvin}.

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable with high availability in the
presence of partitions~\cite{davidson-survey}. However, database
systems have a long tradition of providing weaker isolation and
consistency guarantees~\cite{adya, ansicritique, gray-virtues,
  gray-isolation, kemme-thesis}. Today's ACID and NewSQL databases
often employ weak isolation models due to concurrency and performance
benefits; weak isolation is overwhelmingly the default setting in
these stores and is often the only option offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics provided by current highly available systems. This raises a
natural question: which of these semantics can be provided with high
availability?

To date, the relationship between ACID semantics and high
availability has not been well explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, krikellas-bargain, calvin, kemme-classification} or snapshot
isolation~\cite{daudjee-snapshot,generalizedsnapshot, kemme-snapshot,
  walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

Our main contributions in this paper are as follows. We fill the gap
between the many previously proposed isolation and consistency models
and the goal of high availability.  We classify which among the wide
array of models are achievable with high availability, denoting them
as {\em Highly Available Transactions} (HATs). In doing so, we
demonstrate that although many implementations of HAT semantics are
not highly available, this is an artifact of the implementations
rather than an inherent property of the semantics. Our investigation
shows that, besides serializability, Snapshot Isolation and Repeatable
Read isolation are not HAT-compliant, while most other isolation
levels are achievable with high availability. We also demonstrate that
many weak replica consistency models from distributed systems are both
HAT-compliant and simultaneously achievable with several ACID
properties.

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we show
is unavailable. However, Read Committed isolation, Transactional
Atomicity, and many of the weaker data consistency models from
database and distributed systems are achievable via algorithms that
rely on multi-versioning and limited client-side caching. For several
guarantees, such as causal consistency with phantom prevention and
ANSI Repeatable Read, we consider a modified form of high availability
in which clients ``stick to'' (i.e., have affinity with) at least one
server---a property which is often implicit in the distributed systems
literature~\cite{herlihy-art, eiger, cac} but which requires explicit
consideration in a client-server replicated database context. This
sticky availability is widely employed~\cite{eiger, vogels-defs} but
is less restrictive (and therefore more easily achievable) than
traditional high availability.

Our results demonstrate that a wide range of semantic guarantees can
be delivered with availability in face of partitions, but which
guarantees are worthwhile in practice? We study the virtues and
limitations of both classic and HAT systems by surveying practitioner
accounts and research literature, performing experimental analysis on
modern cloud infrastructure, and analyzing representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic, HATs can fall short, requiring
sometimes-unavailable mechanisms that are more coordination
intensive. Overall, this paper quantifies the benefits that Highly
Available Transactions offer as well as the necessary restrictions
that they place on applications.

%% In this paper, we make the following contributions:
%% \begin{myitemize}
%% \item We model high availability in a transactional environment,
%%   including traditional and sticky high availability.

%% \item We taxonomize ACID and distributed consistency properties
%%   according to their availability characteristics.

%% \item We analyze existing concurrency control algorithms, perform a
%%   case-study of an existing transactional application, and
%%   briefly evaluate a HAT database prototype.
%% \end{myitemize}