### 2013-07-22 writeup/hat-analysis.tex

\section{Highly Available Transactions}
\label{sec:hats}

HAT systems provide transactions with transactional availability and
either high availability or sticky high availability. They offer
latency and availability benefits over traditional distributed
databases, yet they cannot achieve all possible semantics. In this
section, we delineate ACID, distributed replica consistency, and
session consistency levels which can be achieved with high
availability (transactional atomicity, variants of Repeatable Read and
traditionally weaker isolation, and many session guarantees), those
with sticky high availability (read your writes, PRAM and causal
consistency), and properties that cannot be provided in a HAT system
(those preventing Lost Update and Write Skew or guaranteeing recency).
We present a full summary of these results in
Section~\ref{sec:hat-summary}.

As Brewer states, ``systems and database communities are separate but
overlapping (with distinct vocabulary)''~\cite{brewer-slides}. With
this challenge in mind, we build on existing properties and
definitions from the database and distributed systems literature,
providing a brief, informal explanation and example for each
guarantee. The database isolation guarantees require particular care,
since different DBMSs often use the same terminology for different
mechanisms and may provide additional guarantees in addition to our
implementation-agnostic definitions.  We draw largely on Atul Adya's
dissertation~\cite{adya} and somewhat on its predecessor work: the
ANSI SQL specification~\cite{ansi-sql} and Berenson et al.'s
subsequent critique~\cite{ansicritique}.

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in
 Appendix A.
% our extended Technical Report~\cite{hat-tr}.
In our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

\subsection{Achievable HAT Semantics}

To begin, we present well-known semantics that can be achieved in HAT
systems. In this section, our primary goal is feasibility, not
performance. As a result, we offer proof-of-concept highly available
algorithms that are not necessarily optimal or even efficient: the
challenge is to prove the existence of algorithms that provide high
availability. However, we briefly study a subset of their performance
implications in Section~\ref{sec:evaluation}.

\subsubsection{ACID Isolation Guarantees}
\label{sec:isolation}

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times but should
handle concurrent updates (i.e., overwrites) in accordance with the
total order for each item. \textit{PL-1} requires that writes to
different objects be ordered consistently across transactions,
prohibiting Adya's phenomenon $G0$ (also called ``Dirty
Writes''~\cite{ansicritique}). If we build a graph of transactions
with edges from one transaction to another if the former overwrites
the latter's write to the same object, then, under Read Uncommitted,
the graph should not contain cycles~\cite{adya}. Consider the
following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$. Read Uncommitted is easily
achieved via timestamping each update in a transaction with the same
timestamp (unique across transactions; e.g., combining a client's ID
with a sequence number) and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs (as in
Section~\ref{sec:modernacid}). Centralized implementations differ,
with some based on long-duration exclusive locks and short-duration
read locks~\cite{gray-isolation} and others based on multiple
versions. These implementations often provide recency and monotonicity
properties beyond what is implied by the name ``Read Committed'' and
what is captured by the implementation-agnostic definition: under
Read Committed, transactions should not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's \textit{PL-2} and is formalized by prohibiting
Adya's \textit{G1\{a-c\}} (or ANSI's $P1$, or ``broad'' $P1$ [2.2]
from Berenson et al.). For instance, in the example below, $T_3$
should never see $a=1$, and, if $T_2$ aborts, $T_3$ should not read
$a=3$:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\vspace{-1em}
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
client never writes uncommitted data to shared copies of data, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. Unlike a lock-based implementation, this implementation
does not provide recency or monotonicity guarantees but it satisfies
the implementation-agnostic definition.

Several different properties have been labeled \textbf{Repeatable
  Read} isolation. As we will show in
Section~\ref{sec:unachievable-acid}, some of these are not achievable
in a HAT system. However, the ANSI standardized
implementation-agnostic definition~\cite{ansi-sql} \textit{is}
achievable and directly captures the spirit of the term: if a
transaction reads the same data twice, it sees the same value each
time (preventing ``Fuzzy Read,'' or $P2$). In this paper, to
disambiguate between other definitions of ``Repeatable Read,'' we will
call this property ``cut isolation,'' since each transaction reads
from a non-changing cut, or snapshot, over the data items. If this
property holds over reads from discrete data items, we call it
\textbf{Item Cut Isolation}, and, if we also expect a cut over
predicate-based reads (e.g., \texttt{SELECT WHERE}; preventing
Phantoms~\cite{gray-isolation}, or Berenson et al.'s $P3/A3$), we have
the stronger property of \textbf{Predicate Cut-Isolation}. In the
example below, under both levels of cut isolation, $T_3$ must read
$a=1$:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: w_x(2)
\\T_3 &: r_x(1)~r_x(a)
\end{align*}
It is possible to satisfy Item Cut Isolation with high availability by
having transactions cache their reads at the client such that the
values that they read for each item never changes unless they
overwrite it themselves. The cache can be cleared at the end of each
transaction and can alternatively be accomplished on (sticky) replicas
via multi-versioning. Predicate Cut Isolation is also achievable in
HAT systems via similar caching or multi-versioning that track entire
logical ranges of predicates in addition to item-based reads.

\subsubsection{ACID Atomicity Guarantees}
\label{sec:ta}

Atomicity, informally guaranteeing that either all or none of
transactions' effects should succeed, is core to ACID
guarantees. Although, at least by the ACID acronym, it is not an
``isolation'' property, atomicity properties also restrict the updates
visible to other transactions. Accordingly, here, we consider the
\textit{isolation} effects of atomicity, which we call \textbf{Read
  Atomic (RA)} isolation.  Under RA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter, all effects of $T_i$ are observed by $T_j$. That is, if a
transaction $T_j$ reads a version of an object that transaction $T_i$
wrote, then a later read by $T_j$ cannot return value whose next
version is installed by $T_i$. Together with item cut isolation, RA
prevents Read Skew anomalies (Berenson et al.'s A5A) and is useful in
several contexts such as maintaining foreign key constraints,
consistent global secondary indexing, and maintenance of derived
data. In the example below, under RA, because $T_2$ has read $T_1$'s
write to $y$, $T_2$ must observe $b=c=1$ (or later versions for each
key):
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of $x$. In
the hierarchy of existing isolation properties, we place RA below
Adya's $PL-2L$ (as it does not necessarily enforce transitive
read-write dependencies) but above Read Committed ($PL-2$). Notably,
RA requires disallows reading intermediate writes (Adya's $G1b$):
observing all effects of a transaction implicitly requires observing
the final (committed) effects of the transaction as well.

Perplexingly, discussions of RA are absent from existing treatments of
weak isolation. This is perhaps again due to the single-node context
in which prior work was developed: on a single server (or a fully
replicated database), RA is achievable via lightweight locking and/or
local concurrency control over data items~\cite{gstore,
  kemme-thesis}. In contrast, in a distributed environment, RA over
arbitrary groups of non-colocated items is considerably more difficult
to achieve with high availability: servers need to ensure that,
regardless of replica selection, all of a client's accesses will read
a sufficiently up-to-date/``consistent'' set of data items. However,
replicas do not need to agree on when to choose to reveal sets of
values to clients, which would require consensus. Instead, as we see
below, RA only requires the equivalent of reliable broadcast (with
some additional client metadata) and is achievable in HAT systems.

% RA can be achieved via multi-versioning. If clients attach vector
% clocks to every write (incrementing their own position for each
% transaction) and replicas store all writes via multi-versioning, then
% clients can safely determine which sets of data items to read; this
% approach is adopted by Swift~\cite{swift} and, to a lesser-extent,
% bolt-on causal consistency~\cite{bolton}. These systems also provide
% causal consistency and are subsequently sticky available (each is
% implemented via client-side caching). However, RA is achievable
% without stickiness via an alternate algorithm: if servers wait to
% reveal writes until they are present on all replicas, clients do not
% need to be sticky.

As a straw-man, replicas can store all versions ever written for all
items. Servers can gossip information about versions they have seen
and construct a lower bound on the versions that can be found on every
replica. At the start of each transaction, clients choose a bound that
is guaranteed to be lower than or equal to the this global lower
bound, and, during transaction execution, replicas return the latest
version of each item that is not greater than the client's chosen
bound. If the lower bound is advanced along transactional boundaries,
clients will observe RA. This algorithm has several variants in the
literature~\cite{readonly, swift}: bounds can be represented by vector
clocks, and older versions can be asynchronously garbage collected.

We have developed a more efficient RA algorithm, which we sketch here
and provide greater detail in Appendix B. We begin with our Read
Committed algorithm, but replicas wait to reveal new writes to readers
until all of the replicas for the final writes in the transaction have
received their respective writes (are \textit{pending
  stable}). Clients include additional metadata with each write: a
single timestamp for all writes in the transaction (e.g., as in Read
Uncommitted) and a list of items written to in the transaction. When a
client reads, the return value's timestamp and list of items form a
lower bound on the versions that the client should read for the other
items. When a client reads, it attaches a timestamp to its request
representing the current lower bound for that item. Replicas use this
timestamp to respond with either a write matching the timestamp or a
pending stable write with a higher timestamp. Servers keep two sets of
writes for each data item: the write with the highest timestamp that
is pending stable and a set of writes that are not yet pending
stable. This is entirely master-less and operations never block due to
replica coordination. Moreover, there are several possible
optimizations that, for brevity, we do not describe here.

However, one key property with RA is that it do not dictate when
writes become visible; accordingly, the algorithm is highly
available. We can optimize write visibility via sticky availability by
sticking clients are with disjoint groups of servers: once all of a
transaction's writes are pending stable within a group of servers,
writes can be made visible (as opposed to once they are pending stable
with respect to all servers). For example, in a multi-datacenter
setting, all clients and servers within a datacenter may form a sticky
group.

\subsubsection{Session Guarantees}

As is typical in the database literature (with few
exceptions~\cite{daudjee-session}), our models have not yet considered
real-time or client-centric ordering \textit{across} transactions. In
the distributed systems literature~\cite{sessionguarantees, vogels-defs}, many
useful \textit{safety} guarantees span multiple, non-transactional
operations. In particular, \textit{session guarantees} are used to
describe guarantees across transactions within a given
\textit{session}, ``an abstraction for the sequence of...operations
performed during the execution of an
application''~\cite{sessionguarantees}. Informally a session describes
a context that should persist between transactions: for example, on a
social networking site, all of a user's transactions submitted between
``log in'' and ``log out'' operations might form a session.

Several session guarantees can be made with high availability:

\vspace{.5em}\noindent\textbf{{Monotonic reads}} requires that, within
a session, subsequent reads to a given object ``never return any
previous values''; reads from each item progress according to a total
order (e.g., the order from Read Uncommitted).

\vspace{.5em}\noindent\textbf{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted by
the client. Any order on transactions should also be consistent with
any precedence that a global observer would see.

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s); this corresponds to Lamport's
``happens-before'' relation~\cite{lamportclocks}.  Any order on
transactions should respect the transitive reads-from
order.\vspace{.5em}

The above guarantees can be achieved by forcing servers to wait to
reveal new writes (say, by buffering them in separate local storage)
until each write's respective dependencies are visible on all
replicas. This mechanism effectively ensures that all clients read
from a globally agreed upon lower bound on the versions written. This
is highly available as a client will never block due to inability to
find a server with a sufficiently up-to-date version of a data
item. However, it does not imply that transactions will read their own
writes or, in the presence of partitions, make forward progress
through the version history. The problem is that under non-sticky high
availability, a system must handle the possibility that, under a
partition, an unfortunate client will be forced to issue its next
requests against a partitioned, out-of-date server.

The solution to this conundrum is to give up high availability and
settle for sticky availability. Sticky availability permits three
additional guarantees, which we first define and then prove are
unachievable in a generic highly available system:

\vspace{.5em}\noindent\textbf{{Read your writes}} requires that
whenever a client reads a given data item after updating it, the read
returns the updated value (or a value that overwrote the previously
written value).

\vspace{.5em}\noindent\textbf{{PRAM}} (Pipelined Random Access Memory)
provides the illusion of serializing each of the operations (both
reads and writes) within each session and is the combination of
monotonic reads, monotonic writes, and read your
writes~\cite{herlihy-art}.

\vspace{.5em}\noindent\textbf{{Causal
    consistency}}~\cite{causalmemory} is the combination of all of the
session guarantees~\cite{sessiontocausal} (alternatively, PRAM with
writes-follow-reads) and is also referred to by Adya as \textit{PL-2L}
isolation~\cite{adya}).\vspace{.5em}

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, then, for transactional availability,
the server must allow $T_1$ to commit. If the same client subsequently
executes $T_2$ against the same (partitioned) server in the same
session, then it will be able to read its writes. However, if the
network topology changes and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the system will have to either stall indefinitely to allow the client
to read her writes (violating transactional availability) or will have
to sacrifice read your writes guarantees. However, if the client
remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, cops, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier: only reveal new writes to
clients when their dependencies have been revealed.

\subsubsection{Additional HAT Guarantees}

In this section, we briefly discuss two additional kinds of guarantees
that are achievable in HAT systems.

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
application-level consistency guarantees. It can often execute
commutative and logically monotonic~\cite{calm} operations without the
risk of invalidating application-level integrity constraints and can
maintain limited criteria like foreign key constraints (via TA). We do
not attempt to describe the entire space of application-level
consistency properties that are achievable (see
Section~\ref{sec:relatedwork}) but we specifically evaluate TPC-C
transaction semantics with HAT guarantees in
Section~\ref{sec:evaluation}.

\vspace{.5em}\noindent{\textbf{Convergence}} To require that the
system propagates writes between replicas, we can require convergence,
or \textit{eventual consistency}: in the absence of new mutations to a
data item, in the absence of partitions, all servers should eventually
agree on the value for each item~\cite{cac, vogels-defs}. This is
typically accomplished by any number of anti-entropy protocols, which
periodically update neighboring servers with the latest value for each
data item~\cite{antientropy}. Establishing a final convergent value is
related to determining a total order on transaction updates to each
item, as in Read Uncommitted.

\subsection{Unachievable HAT Semantics}
\label{sec:unachievable-hat}

While there are infinitely many HAT models
(Section~\ref{sec:relatedwork}), at this point, we have largely
exhausted the range of achievable, previously defined (and useful)
semantics that are available to HAT systems. Before summarizing our
possibility results, we will present impossibility results for HATs,
also defined in terms of previously identified isolation and
consistency anomalies. Most notably, it is impossible to
prevent Lost Update or Write Skew in a HAT system.

\subsubsection{Unachievable ACID Isolation}
\label{sec:unachievable-acid}

In this section, we demonstrate that preventing Lost Update and Write
Skew---and therefore providing Snapshot Isolation, Repeatable Read,
and one-copy serializability---inherently requires foregoing high
availability guarantees.

Berenson et al. define \textit{Lost Update} as when one
transaction $T1$ reads a given data item, a second transaction $T2$
updates the same data item, then $T1$ modifies the data item based on
its original read of the data item, ``missing'' or ``losing'' $T2$'s
newer update. Consider a database containing only the following
transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: r_x(a)~w_x(a+2)
\\T_2 &: w_x(2)\vspace{-1em}
\end{align*}
If $T_1$ reads $a=1$ but $T_2$'s write to $x$ precedes $T_1$'s write
operation, then the database will end up with $a=3$, a state that
could not have resulted in a serial execution due to $T_2$'s
``Lost Update.''

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small\vspace{-1em}
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)\vspace{-1em}
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen from a serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but the problem persists in the
  general case.}  To prevent this, either $T_1$ or
$T_2$ should not have committed. Each client's respective server might
try to detect that another write occurred, but this requires knowing
the version of the latest write to $x$. In our example, this reduces
to a requirement for linearizability, which is, via Gilbert and
Lynch's proof of the CAP Theorem, provably at odds with high
availability~\cite{gilbert-cap}.

\textbf{Write Skew} is a generalization of Lost Update to multiple
keys. It occurs when one transaction $T1$ reads a given data item $x$,
a second transaction $T2$ reads a different data item $y$, then $T1$
writes to $y$ and commits and $T2$ writes to $x$ and commits. As an
example of Write Skew, consider the following two transactions:
\begin{align*}
\small
T_1 &: r_y(0)~w_x(1)
\\T_2 &: r_x(0)~w_y(1)
\end{align*}
As Berenson et al. describe, if there was an integrity constraint
between $x$ and $y$ such that only one of $x$ or $y$ should have value
$1$ at any given time, then this write skew would violate the constraint (which is preserved in serializable executions). Write skew is a somewhat
esoteric anomaly---for example, it does not appear in
TPC-C~\cite{snapshot-serializable}---but, as a generalization of Lost
Update, it is also unavailable to HAT systems.

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation (including Parallel Snapshot Isolation~\cite{walter}), and
Cursor Stability guarantees are all unavailable.  Repeatable Read
(defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability~\cite{1sr} need to prevent both Lost Update and Write Skew. These
prevention requirements mean that these guarantees are inherently
unavailable.

\subsubsection{Unavailable Recency Guarantees}

Distributed data storage systems often make various recency guarantees
on reads of data items.  Unfortunately, an indefinitely long partition
can force an available system to violate any recency bound, so recency
bounds are not enforceable by HAT systems~\cite{gilbert-cap}. One of
the most famous of these guarantees is
linearizability~\cite{herlihy-art}, which states that reads will
return the last completed write to a data item, and there are several
other (weaker) variants such as safe and regular register
semantics. When applied to transactional semantics, the combination of
one-copy serializability and linearizability is called \textit{strong
  (or strict) one-copy serializability}~\cite{adya} (e.g.,
Spanner~\cite{spanner}). It is also common, particularly in systems
that allow reading from masters and slaves, to provide a guarantee
such as ``read a version that is no more than five seconds out of
date'' or similar. None of these guarantees is HAT-compliant.

\subsubsection{Durability}

A client requiring that its transactions' effects survive $F$ server
faults requires at least $F+1$ non-failing replicas; this affects
availability and, according to the Gilbert and Lynch definition we
have adopted, $F>1$ fault tolerance is not achievable with high
availability.

\subsection{Summary}
\label{sec:hat-summary}

As we summarize in Table~\ref{table:hatcompared}, a wide range of
isolation levels are achievable in HAT systems, including
transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve read your
writes guarantees and PRAM and causal consistency. However, many other
prominent semantics, such as Snapshot Isolation, One-Copy
Serializability, and Strict Serializability cannot be achieved due to
the inability to prevent Lost Update and Write Skew phenomena.

We illustrate the hierarchy of available, sticky available, and
unavailable consistency models we have discussed in
Figure~\ref{fig:hatcompared}. Many models are simultaneously
achievable, but we find several particularly compelling. If we combine
all sticky-HAT guarantees, we have transactional, causal snapshot
reads (i.e., Causal Transactional Predicate Cut Isolation). If we
combine RA and P-CI, we have transactional snapshot reads. We can
achieve RC, MR, and RYW by simply sticking clients to servers. We can
also combine unavailable models---for example, an unavailable system
might provide PRAM and One-Copy
Serializability~\cite{daudjee-session}.

To the best of our knowledge, this is the first unification of
database ACID transactions, distributed consistency, and session
guarantee models. Interestingly, strong one-copy serializability
entails all other models, while considering the (large) power set of
all compatible models (e.g., the diagram depicts 144 possible HAT
combinations) hints at the vast expanse of consistency models found in
the literature. This taxonomy is not exhaustive
(Section~\ref{sec:conclusion}), but we believe it lends substantial
clarity to the relationships between a large subset of the prominent
ACID and distributed consistency models. Additional read/write
transaction semantics that we have omitted should be classifiable
based on the available primitives and HAT-incompatible anomaly
prevention we have already discussed.

 \newcommand{\lostupdate}{$^\dagger$}
 \newcommand{\rwskew}{$^\ddagger$}
 \newcommand{\linearizable}{$^\oplus$}

\begin{table}[t!]
\begin{tabular}{| c | p{6cm} | }\hline
HA & Read Uncommitted (RU), Read Committed (RC), Read
Atomic (RA), Item Cut Isolation (I-CI), Predicate Cut Isolation
(P-CI), Writes Follow Reads (WFR), Monotonic Reads (MR), Monotonic
Writes (MW)\\\hline Sticky HA & Read Your Writes (RYW), PRAM,
Causal\\\hline Unavailable & Cursor Stability (CS)\lostupdate,
Snapshot Isolation (SI)\lostupdate, Repeatable Read
(RR)\lostupdate\rwskew, One-Copy Serializability
(1SR)\lostupdate\rwskew, Recency\linearizable, Safe\linearizable,
Regular\linearizable, Linearizability\linearizable, Strong
1SR\lostupdate\rwskew\linearizable \\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing write skew\rwskew, and requiring
  recency guarantees\linearizable.}
\label{table:hatcompared}
\end{table}

\begin{figure}[t!]
\centering
\begin{tikzpicture}[scale=0.8]
  \tikzstyle{sticky}=[rectangle,draw=blue!50,fill=blue!20,thick]
  \tikzstyle{noha}=[ellipse,draw=red!50,fill=red!20,thick, inner sep=0pt,minimum size=12pt]

  \tikzstyle{every node}=[font=\small]

 \node[draw=none,fill=none] (ici) at (1.2, 0) {I-CI};
 \node[draw=none,fill=none] (pci) at (1.75, .95) {P-CI};
 \node[draw=none,fill=none] (rc) at (-1.2, .95) {RC};
 \node[draw=none,fill=none] (ru) at (-1.2, 0) {RU};

 \node[draw=none,fill=none] (ra) at (0, 1.475) {RA};

 \node[draw=none,fill=none] (mr) at (3.6, 0) {MR};
 \node[draw=none,fill=none] (mw) at (4.8, 0) {MW};
 \node[draw=none,fill=none] (wfr) at (2.4,0) {WFR};
 \node at (6.1,0) [sticky] (ryw) {RYW};

 \node[noha](recency) at (7.7, 0) {recency};
 \node[noha](safe) at (7.7, 1) {safe};
 \node[noha](regular) at (7.7, 2) {regular};
 \node[noha](linearizable) at (7.7, 3) {linearizable};
 \node at (4.8, 2) [sticky] (causal) {causal};
 \node at (4.8, 1) [sticky] (pram) {PRAM};
 \node[noha] (cs) at (-1.2, 2) {CS};
 \node[noha] (rr) at (0.2, 2.7) {RR};
 \node[noha] (si) at (1.75, 2) {SI};
 \node[noha] (1sr) at (1.75, 3.2) {1SR};
 \node[noha] (ssr) at (3.85, 3.6) {Strong-1SR};

 \draw [->, red] (recency) -- (safe);
 \draw [->, red] (safe) -- (regular);
 \draw [->, red] (regular) -- (linearizable);
 \draw [->, red] (linearizable) -- (ssr);
 \draw [->, red] (1sr) -- (ssr);

 \draw [->] (ru) -- (rc);
 \draw [->] (rc) -- (ra);
 \draw [->] (ici) -- (pci);

 \draw [->, blue] (mr) -- (pram);
 \draw [->, blue] (mw) -- (pram);
 \draw [->, blue] (wfr) -- (causal);
 \draw [->, blue] (ryw) -- (pram);
 \draw [->, blue] (pram) -- (causal);

 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ru) -- (mr);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (rc) -- (ta);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ici) -- (ta);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (pci) -- (ta);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (rc) -- (mr);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (pci) -- (mr);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ici) -- (mr);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ta) -- (ru);
 %\draw[snake=coil, blue, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ru) -- (causal);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (mr) -- (mw);
 %\draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (wfr) -- (mw);
 %\draw[snake=coil, blue, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (wfr) -- (ryw);

 \draw [->, red] (rc) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (pci) -- (si);
 \draw [->, red] (ici) -- (rr);
 \draw [->, red] (rr) -- (1sr);
 \draw [->, red] (si) -- (1sr);
 \draw [->, red] (ra) -- (si);
 \draw [->, red] (ra) -- (rr);
 \draw [->, red] (causal) -- (linearizable);
 \draw [->, red] (ryw) -- (safe);

\end{tikzpicture}
\label{fig:hat-order}
\caption{Partial ordering of HAT, sticky HAT (in boxes, blue), and
  unavailable models (circled, red) from
  Table~\protect\ref{table:hatcompared}. Directed edges represent
  ordering by model strength. Incomparable models can be
  simultaneously achieved, and the availability of a combination of
  models has the availability of the least available individual
  model.}\vspace{-1em}
\label{fig:hatcompared}
\end{figure}

\subsection{Additional Discussion}
\label{sec:discussion}

In this section, we discuss several subtleties in our results,
specifically addressing model composition, transactional atomicity
versus linearizability, and benefits of stickiness.

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-1em}
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and the system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs
to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute (e.g.,
COPS~\cite{cops}). Using client-side caching can alleviate some of
these challenges~\cite{bolton, swift}, but then the system becomes
sticky high available.

Composition cost may also vary by combination. For instance, Charron-Bost
 proved that, to capture causality between $N$ communicating
processes, standard vector-based approaches face an upper bound of
$O(N)$ storage per write~\cite{charron-bost}. This means that, with
$100K$ clients, each write might be accompanied by $100K$ timestamps
per vector. This is difficult to scale. By compromising on
availability (e.g., treating a datacenter as a linearizable cluster),
this overhead can be reduced~\cite{eiger}, but it is much
cheaper to provide, say, read your writes guarantees, than causal
consistency.

\vspace{.5em}\noindent\textbf{Linearizability and Transactional
  Atomicity} The relationship between linearizability and
transactional atomicity is non-obvious. RA dictates that writes to
multiple keys across multiple servers are made visible to readers all
at once, while linearizability dictates that writes to a single key on
multiple servers are made visible to all readers at once---what is
different? First, in linearizable (and safe and regular) systems,
writes are made visible to clients \textit{immediately} after they
finish. With transactional atomicity, there is no recency
guarantee. Second, in linearizable systems, all clients see all writes
at the same time. With RA as defined here, clients may see writes at
different times depending on which replicas they contact. We are not
aware of an analogous model in the distributed systems
literature. Accordingly, despite apparent similarities, RA is
incomparable with and less coordination intensive (by availability
measures) than linearizability.

\vspace{.5em}\noindent\textbf{Visibility and Stickiness} Sticky
availability can improve write \textit{visibility}: clients will be
able to safely read writes more quickly in a sticky available
system. In the model we discussed, it is possible to achieve several
properties like monotonic reads in a highly available system by
waiting to reveal a write until all servers have seen it and its
relevant dependencies. However, this incurs severe visibility
penalties---new writes will not become visible to clients in the
presence of partitions. A client that does not want to guarantee
read-your-writes (due to the sticky availability requirement) may
still wish to read other clients' writes with timeliness.