### 2013-10-01 writeup/conclusion.tex
Editing pass

\section{Conclusions and Future Work}
\label{sec:conclusion}

The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---``always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are
not. Surprisingly, many of the default (and sometimes strongest)
semantics provided by today's traditional database systems are
achievable as HATs, hinting that distributed databases need not
compromise availability in order to serve many existing applications.

HATs raise several challenges for future study. In this paper, we have
largely focused on previously defined isolation and data consistency
models from the consistency from the database and distributed systems
communities. Their previous definitions and, in many cases, widespread
adoption hints at their utility to end-users. However, we believe
there is considerable work to be done to improve the programmability
of highly available systems. Isolation and data consistency are means
by which \textit{application-level} consistency is achieved but are
typically not end goals for end-user applications. Our results hint
that a mix of HAT and non-HAT semantics (the latter used sparingly) is
required for practical applications, but the decision to employ each
and the system architectures for a hybrid approach remain open
problems. While we have studied the analytical and experiment behavior
of several HAT models, there is substantial work in further
understanding the performance and design of systems within the large
set of HAT models. Weakened failure assumptions as in escrow or in the
form of bounded network asynchrony could enable richer HAT semantics
at the cost of general-purpose availability. Alternatively, there is a
range of possible solutions providing strong semantics during
partition-free periods and weakened semantics during partitions. Based
on our understanding of what is desired in the field and newfound
knowledge of what is possible to achieve, we believe HATs provide a
large and useful design space for exploration.

%% In this paper we expose a broad design space of Highly Available Transactional 
%% Systems (HATs), which can offer the key benefits of
%% highly available distributed systems---partition tolerance and low-latency 
%% writes---while also providing a family of transactional isolation 
%% levels and replica semantics that have been widely used in practice. 
%% In taxonomizing this design space, we show that many transactional 
%% guarantees are achievable with high availability, while a few, 
%% like preventing Lost Update and Write Skew, are not.

%% There has been some debate in the database community regarding the importance
%% of availability in modern databases; there is also concern about the utility of
%% relaxed isolation.  Our analysis and experiments in Sections~\ref{sec:motivation} 
%% and~\ref{sec:evaluation} convince us that the availability goals of the NoSQL
%% movement will remain a critical aspect of many important systems going forward, and 
%% that the isolation levels available via HATs can have significant practical impact
%% in delivering useful transactional semantics to application developers.

%% Much of the prior research in distributed transactions has focused on prototyping 
%% individual points in the design space, with a focus on providing serializability
%% at the expense of availability.  Our goal in this paper has been to take a broader view. 
%% We taxonomize the joint design space of replica consistency and transactional isolation,
%% and identify what is possible and impossible to achieve semantically via HATs.  From
%% a research perspective, this offers a long-overdue connection between storage 
%% semantics studied in the database and distributed systems communities.
%% With this understanding in hand, the next challenge is to return to issues of 
%% performance and systems architecture within the HATs design space.  Based on our 
%% understanding of what is desired in the field and what is possible to achieve, we 
%% believe this is an area of great opportunity in the coming years.