### 2013-10-01 writeup/hat-vldb.tex

% THIS IS AN EXAMPLE DOCUMENT FOR VLDB 2012
% based on ACM SIGPROC-SP.TEX VERSION 2.7
% Modified by  Gerald Weber <gerald@cs.auckland.ac.nz>
% Removed the requirement to include *bbl file in here. (AhmetSacan, Sep2012)
% Fixed the equation on page 3 to prevent line overflow. (AhmetSacan, Sep2012)

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}
\usepackage[hyphens]{url}

\usepackage[hidelinks]{hyperref}

\usepackage{breakurl}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}}

\newif\iftechreport
%\techreporttrue
\techreportfalse

\interfootnotelinepenalty=10000

\usepackage{times, graphicx, xcolor, multicol, minipage, subfig, tikz, amsthm, float}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\newcommand{\subparagraph}{}

\usepackage{xspace, titlesec}

\newcommand{\etal}{{\em et al.}\xspace}
\newcommand{\etc}{{\em etc}\xspace}
\newcommand{\cf}{{\em c.f.,}\xspace}
\newcommand{\versus}{{\em vs.}\xspace}
\newcommand{\ie}{{\em i.e.,}\xspace}
\newcommand{\eg}{{\em e.g.,}\xspace}

\newcommand{\figfactor}{1.0}

\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx
\newcommand{\commentt}[1]{{\small\texttt{#1}}}

\theoremstyle{definition}
\newtheorem{observation}{Observation}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

\renewcommand{\ttdefault}{fi4}

\usetikzlibrary{positioning, snakes, shapes}

\tikzset{>=latex}

\titleformat*{\section}{\fontsize{12pt}{12}\selectfont\bfseries}
\titleformat*{\subsection}{\fontsize{10pt}{11}\selectfont\bfseries}
\titleformat*{\subsubsection}{\fontsize{10pt}{11}\itshape}

\titlespacing{\section}{1ex}{1ex}{0.25ex}

\titlespacing{\subsection}{1ex}{0.5ex}{0.25ex}
\titlespacing{\subsubsection}{1ex}{0.25ex}{0.25ex}

\let\footnotesize\small

\newcommand*{\DashedArrow}[1][]{\mathbin{\tikz [baseline=-.25ex,-latex, dashed,#1] \draw [#1] (0pt,0.5ex) -- (2em,0.5ex);}}%

\newenvironment{myitemize}
{

   \vspace{0mm}
    \begin{list}{$\bullet$}{\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

\newenvironment{myenumerate}
{

   \vspace{0mm}
   \newcounter{qcounter}
    \begin{list}{\arabic{qcounter}.~}{\usecounter{qcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

\newenvironment{introenumerate}
{

   \vspace{-.5em}
   \newcounter{qdcounter}
    \begin{list}{\arabic{qdcounter}.~}{\usecounter{qdcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

\begin{document}

% ****************** TITLE ****************************************

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis, Aaron Davidson, Alan Fekete{\fontsize{12}{14}$^\dagger$}, Ali Ghodsi, Joseph M. Hellerstein, Ion Stoica \\{\affaddr{UC Berkeley and {\fontsize{12}{14}$^\dagger$}University of Sydney}}}}
\maketitle

\begin{abstract}
\noindent To minimize network latency and remain online during server
failures and network partitions, many modern distributed data storage
systems eschew transactional functionality, which provides strong
semantic guarantees for groups of multiple operations over multiple
data items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak transactional isolation, replica consistency,
and highly available systems. We analytically and experimentally
quantify the availability and performance benefits of HATs---often two
to three orders of magnitude over wide-area networks---and discuss
their necessary semantic compromises.
\end{abstract}

\input{intro.tex}

\input{motivation.tex}

\input{model.tex}

\input{hat-analysis.tex}

\input{evaluation.tex}

\input{relatedwork.tex}

%\input{futurework.tex}

\input{conclusion.tex}

\section*{Acknowledgments}
We would like to thank Peter Alvaro, Neil Conway, Evan Jones, Adam
Oliner, Aurojit Panda, and Shivaram Venkataraman for their helpful
feedback on this work. This research was supported in part by the Air
Force Office of Scientiﬁc Research (grant FA95500810352), DARPA XData
Award FA8750-12-2-0331, the Natural Sciences and Engineering Research
Council of Canada, the National Science Foundation (grants
CNS-0722077, IIS-0713661, and IIS-0803690), NSF CISE Expeditions award
CCF-1139158, the National Science Foundation Graduate Research
Fellowship (grant DGE-1106400), and gifts from Amazon Web Services,
Google, SAP, Cisco, Clearstory Data, Cloudera, Ericsson, Facebook,
FitWave, General Electric, Hortonworks, Huawei, Intel, Microsoft,
NetApp, Oracle, Samsung, Splunk, VMware, WANdisco and Yahoo!.

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
%\footnotesize
\scriptsize

\bibliography{hat-vldb}

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page

\iftechreport
\pagebreak
\begin{appendix}

\input{hat-appendix.tex}
\end{appendix}
\fi
\end{document}

% megastore
% redblue
% gemini
% redblue
% granola
% bernstein-concurrency
% surviving-failures-bodik
% blooml
% google-availability
% cops