### 2013-10-02 writeup/conclusion.tex
More editing

\section{Conclusions and Future Work}
\label{sec:conclusion}

The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. We expose a broad design space of Highly Available
Transactions (HATs), which can offer the key benefits of highly
available distributed systems---``always on'' operation during
partitions and low-latency operations (often orders of magnitude lower
than non-compliant)---while also providing a family of transactional
isolation levels and replica semantics that have been adopted in
practice.  We also identify many semantic guarantees that are
unachievable with high availability, including Lost Update and Write
Skew anomaly prevention, concurrent update prevention, and bounds on
data recency. Despite these limitations, and somewhat surprisingly,
many of the default (and sometimes strongest) semantics provided by
today's traditional database systems are achievable as HATs, hinting
that distributed databases need not compromise availability, low
latency, or scalability in order to serve many existing applications.

In this paper, we have largely focused on previously defined isolation
and data consistency models from the consistency from the database and
distributed systems communities. Their previous definitions and, in
many cases, widespread adoption hints at their utility to
end-users. However, we believe there is considerable work to be done
to improve the programmability of highly available systems. Isolation
and data consistency are means by which \textit{application-level}
consistency is achieved but are typically not end goals for end-user
applications. Our results hint that a mix of HAT and non-HAT semantics
(the latter used sparingly) is required for practical applications,
but the decision to employ each and the system architectures for a
hybrid approach remain open problems. While we have studied the
analytical and experiment behavior of several HAT models, there is
substantial work in further understanding the performance and design
of systems within the large set of HAT models. Weakened failure
assumptions as in escrow or in the form of bounded network asynchrony
could enable richer HAT semantics at the cost of general-purpose
availability. Alternatively, there is a range of possible solutions
providing strong semantics during partition-free periods and weakened
semantics during partitions. Based on our understanding of what is
desired in the field and newfound knowledge of what is possible to
achieve, we believe HATs represent a large and useful design space for
exploration.

%% In this paper we expose a broad design space of Highly Available Transactional 
%% Systems (HATs), which can offer the key benefits of
%% highly available distributed systems---partition tolerance and low-latency 
%% writes---while also providing a family of transactional isolation 
%% levels and replica semantics that have been widely used in practice. 
%% In taxonomizing this design space, we show that many transactional 
%% guarantees are achievable with high availability, while a few, 
%% like preventing Lost Update and Write Skew, are not.

%% There has been some debate in the database community regarding the importance
%% of availability in modern databases; there is also concern about the utility of
%% relaxed isolation.  Our analysis and experiments in Sections~\ref{sec:motivation} 
%% and~\ref{sec:evaluation} convince us that the availability goals of the NoSQL
%% movement will remain a critical aspect of many important systems going forward, and 
%% that the isolation levels available via HATs can have significant practical impact
%% in delivering useful transactional semantics to application developers.

%% Much of the prior research in distributed transactions has focused on prototyping 
%% individual points in the design space, with a focus on providing serializability
%% at the expense of availability.  Our goal in this paper has been to take a broader view. 
%% We taxonomize the joint design space of replica consistency and transactional isolation,
%% and identify what is possible and impossible to achieve semantically via HATs.  From
%% a research perspective, this offers a long-overdue connection between storage 
%% semantics studied in the database and distributed systems communities.
%% With this understanding in hand, the next challenge is to return to issues of 
%% performance and systems architecture within the HATs design space.  Based on our 
%% understanding of what is desired in the field and what is possible to achieve, we 
%% believe this is an area of great opportunity in the coming years.